---
title: "02 Key R programming concepts in data analysis"
format: html
engine: knitr
---

::: {#exr-01}
**Logical Expressions**

You've got four expressions and two lines of code.
Make both of them to be `TRUE` inputting instead `_____` one of the expressions:

1. `23-39`
2. `sqrt(16)-16`
3. `20-abs(-32)`
4. `-(4^2)`


```{r}
#| eval: false
_____ == ______
_____ == ______
```

:::

---

::: {#exr-02}
**Logical Vectors**

Create a vector `logic_vec` containing logical values for whether each number in `1:10` is **even**.

:::

---

::: {#exr-03}
**Random Numbers & Names**

1. Create a vector `scores` of 7 random numbers between -100 and 100. Use function `runif()`.
2. Assign names to the vector using the days of the week.
3. Find and display all days where the score is **more than -50**.

```{r}
weekdays <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")
```
:::

---

::: {#exr-04}
**Vector Sums**

Combine two randomly created vectors `vec1` and `vec2` of equal length into a new vector `vec_sum`. 
Check whether the sum of `vec_sum` matches the sum of the individual vectors.
:::

---

::: {#exr-05}
**Logic & NA (Part 1)**

You've got a vector with ages:
`ages <- c(12, 25, 8, 34, 16, 40, 10, 21, 18, 30)`

1. Identify people who are **adults** (age 18 and above).
2. Identify **teenagers** (age between 13 and 17, inclusive).
3. Replace ages of **non-adults** with `NA`.

```{r}
#| eval: false
ages <- c(12, 25, 8, 34, 16, 40, 10, 21, 18, 30)

is_adult <- ages ______ 18
is_teenager <- (ages ______ 13) & (ages ______ 17) 
ages[______] <- NA 
```
:::

---

::: {#exr-06}
**Logic & NA (Part 2)**

Continue with the `ages` vector:

1. Identify ages that are `NA` (you can do it with the function `is.na()`).
2. Replace missing values (`NA`) with the **mean** of the remaining ages.
3. Create a logical vector to identify people aged 30 and above (`is_senior`).
4. Count the number of seniors.

```{r}
#| eval: false
is_missing <- ______(ages)
ages[is_missing] <- ______(ages[!is_missing])  
is_senior <- ages ______ 30  
num_seniors <- ______(is_senior) 
```
:::

---

::: {#exr-07}
**Logic Range**

Check if variable `x` is **NOT** between 10 (excluding) and 20 (including).
Make `x` to be equal to 20, then to 15, then to 25 to test your logic.
:::

---

::: {#exr-08}
**Comparing Vectors**

Use a logical vector to identify elements in `vec1` that are either **equal or greater** than their corresponding values in `vec2`.
Display these elements for `vec1`. Do it with a single line of code.

```{r}
#| eval: false
vec1 <- c(10, 20, 30, 40, 50)
vec2 <- c(50, 40, 30, 20, 10)
```
:::

---

::: {#exr-09}
**If Condition**

Write an `IF` condition to check if the sum of elements in `vec1` is positive.
- If so, print "Positive balance".
- Otherwise, print "Negative balance".

```{r}
#| eval: false
vec1 <- c(100, -50, -20)
```
:::

---

::: {#exr-10}
**Condition Chains**

Write an `IF` condition for a variable `temperature`.
Let `temperature <- 30`.

- If it's greater than 25, print "Hot".
- If it's between 10 and 25 (inclusive), print "Moderate".
- Otherwise, print "Cold".

```{r}
#| eval: false
temperature <- 30
```
:::

---

::: {#exr-11}
**Ifelse & Coin Toss**

Simulate a coin toss (1 for heads, 0 for tails).
Use `ifelse()` to assign "Win" for heads and "Lose" for tails in a vector of 1000 random coin tosses.
:::

---

::: {#exr-12}
**Nested Ifelse (BMI)**

Given a vector with BMI, create a new vector using `ifelse()` to classify BMI as:

- "Good" (less than 25)
- "Moderate obesity" (25 to 35 including)
- "Obesity" (more than 35)

```{r}
#| eval: false
bmi <- c(24.4,36.7,38.0,36.9,30.8,34.8,31.8,26.7,36.7,34.9,36.2,24.9,26.6,32.7,24.6,30.8,30.6,38.6,26.2,26.6)
```
:::

---

::: {#exr-13}
**Matrices**

Create a 3x3 matrix with random numbers between 1 and 100 and save to `mat`.

:::

---

::: {#exr-14}
**Matrix Indexing**

Replace all numbers greater than 50 with their half using matrix indexing on `mat`.
:::

---

::: {#exr-15}
**Named Matrices**

Create a named 2x3 matrix representing sales data for two products across three months.
Calculate the total sales for each month and each product.

:::

---

::: {#exr-16}
**Identity Matrix**

1. Create a 4x4 identity matrix. Use function `diag()` for this.
2. Replace its diagonal with the numbers 1, 4, 9, 16.

:::

---

::: {#exr-17}
**Matrix Analysis (TikTok)**

Create a dataset with 5 tiktok users (rows) and 3 weeks (columns) with amount of tiktoks watched.
You may use `runif` again, if you'd like to.

Make a single line of code to print in console the answer to:
1. The week with the highest total tiktoks watched.
2. The user with the highest average tiktoks watched per week.
:::