---
title: "03 Key R programming concepts in data analysis | Part 2"
format: html
engine: knitr
---

::: {#exr-01}
**Factors: Creation**

Create a vector `vec` for outfit ratings: "Casual", "Drip", "Drip", "Casual", "Drip", "Slay", "Drip", "Slay", "Casual", "Drip", "Casual", "Slay".

Convert it into a factor `fact` with levels "Casual", "Drip", "Slay".

```{r}
#| eval: false
vec <- c("Casual", "Drip", "Drip", "Casual", "Drip", "Slay", 
         "Drip", "Slay", "Casual", "Drip", "Casual", "Slay")
# Your code here

```

:::

---

::: {#exr-02}
**Factors: Renaming Levels**

Replace the levels "Casual", "Drip", "Slay" with "C", "D", "S" in the factor you created previously.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-03}
**Factors: Adding Levels (\*)**

Add a new factor level "Cringe" to the factor you created previously and add two elements "Cringe" in the end.

*Hint:* For this, you need to recreate `fact` as character, combine with 2 "Cringe"s and put in levels the levels of previously created fact. Do the addition of elements in a single line of code.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-04}
**Factors: Specific Rename**

Rename only "Cringe" level of the factor to "Cr".

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-05}
**Factors: Summary**

Generate a summary of the factor you created previously.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-06}
**Simulation: Coin Toss**

1. Simulate a coin toss (1 for heads, 0 for tails).
2. Use `ifelse()` to assign "Win" for heads and "Lose" for tails in a vector of 1000 random coin tosses.
3. Make it a factor and then show the summary.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-07}
**Data Frames: Creation**

Create a data frame `city_ua` with the following vectors:

* **City**: "Kyiv", "Lviv", "Odesa", "Kharkiv", "Dnipro"
* **Population** (millions): 2.9, 0.7, 1.0, 1.4, 1.0
* **Subway**: TRUE, FALSE, FALSE, TRUE, TRUE

Extract the Population of "Odesa" in a matrix style.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-08}
**Data Frames: Filtering**

Extract all the rows where the population is greater than 1 million. Do it both in `data.frame` style and in matrix style.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-09}
**Data Frames: Logical Subset**

Create a subset of cities that do not have subway systems. Do it in both data frame style and matrix style.

*Challenge:* To make things funnier, in your line of code don't use `TRUE`, `T`, `FALSE`, `F` :)

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-10}
**Data Frames: New Column (*)**

Add a new column to the data frame indicating whether the city is located in Eastern Ukraine (`TRUE` for "Kharkiv" and "Dnipro", `FALSE` otherwise).

Try the solution without and with the operator `%in%` to check how it works.

> Example of how `%in%` works: `vec %in% c("Cringe","Slay")` will give a logical vector indicating elements that are either "Cringe" or "Slay".

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-11}
**Data Frames: Calculation**

Calculate the mean population of cities that have a subway.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-12}
**Data Analysis: Loading Data**

Load the [`diabetes.csv`](class/diabetes.csv) into `diab` and print the summary of the data. Discuss.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-13}
**Data Analysis: Basic Operations**

Do the following actions with a line of code for each:

1. Find the mean of the BMI column in the dataset.
2. Select all rows where the Age is greater than 50.
3. Select all patients with Age more than 50 and diabetes.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-14}
**Data Analysis: Conditional Factor**

Create a new factor column `bmi_level` in the dataset that indicates whether the BMI is greater than 25 (label as "High BMI" or "Normal BMI").

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-15}
**Data Analysis: Group Comparison**

In the diabetes dataset, compare the average BMI of those who have diabetes and those who don't.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-16}
**Data Analysis: Sorting**

Create sorted diabetes by increasing BMI. Check average Outcome in the first half and in the second half of the dataset. Explain the result.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-17}
**Control Flow: While Loop**

Create a while loop where `i` starts at 26.
Print "i is not divisible by 11" (where i is a current number) in each iteration and stop if the square is divisible by 11.
After the loop print "i is divisible by 11" using final "i".


::: {.callout-tip collapse="true"}
You may want to use `paste0()` function in this task.
:::


```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-18}
**Data Frames: Student Grades**

Create your own dataframe with students that attend these seminars. You need at least 7 students.

Then you need columns:

* **Name** (up to you)
* **Exam I score**, **Exam II score** (put some grades from 0 to 20)
* Column **"passed"**, factor with values "y" or "n". Fill it with "y" if total grade is >=30, "n" otherwise.

Tasks:

1. Reorder dataset from top student to "not that top" by Exam II.
2. Calculate mean and max score for Exam I.
3. Show student names who has top grade in each exam and overall.
4. Adjust dataset values in a way to have three different students in output 🙂

```{r}
#| eval: false
# Your code here

```

:::