---
title: "04 Advanced R programming concepts"
format: html
engine: knitr
---

::: {#exr-01}
**Loops: Cumulative Sum**

1. Create a vector `vec` with 100 elements from 0 to 10 using `runif()`.
2. Write a `for` loop to find the cumulative sum of `vec` and store it in a vector called `cumul`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-02}
**Functions: Basic Creation**

Write a function `cube()` that takes a number and returns its cube ().
Test it on numbers: 4, -2.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-03}
**Functions: Arguments & Logic**

Modify the `cube()` function to allow for a logical argument `show_info` (default `FALSE`).

* If `show_info` is `TRUE`, the function should print: "x to the power of 3 equal to y" (where x and y are actual numbers).
* Test on 4 with `T` and -2 with `F`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-04}
**Lists: Creation**

Create a list called `student` containing three elements:

1. A **name** (character).
2. **Scores** (numeric vector).
3. A boolean **passed** indicating whether the average score is .

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-05}
**Lists: Summary**

Write a function `list_summary()` that takes a list and returns its element names and types (classes).
Test it on the list `student` made previously.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-06}
**Apply Family: lapply**

1. Create a dataset `temp_cities` of city temperatures for five days in five cities.
2. Use `lapply()` to calculate the mean temperature for each city.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-07}
**Apply Family: sapply**

1. Write a function `day_range()` that calculates the temperature range (`max` - `min`) for a vector.
2. Use `sapply()` to apply it to `temp_cities`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-08}
**Functions: Custom Statistics**

Write a function `custom_summary()` that calculates the **mean**, **median**, and **standard deviation** of a numeric vector.
Test it on `vec` (from Exercise 1 or a new one).

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-09}
**Functions: Normalization**

1. Create a function `normalize()` that takes a numeric vector and rescales its values between 0 and 1. Formula: .
2. Use `sapply()` to normalize each city in `temp_cities`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-10}
**Data Import: Read Excel**

Use function `read_excel()` from the library `readxl` to load the [`urbanpop.xlsx`](class/urbanpop.xlsx) file and save it into `pop_1`.

> Note: Don't forget about `setwd()` routine or use the full path to the file.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-11}
**Data Import: Sheets**

1. Use `excel_sheets()` on `urbanpop.xlsx`. Discuss your findings.
2. Try to load the second sheet into `pop_2` using both its **number** (index) and its **name**.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-12}
**Data Import: Bulk Loading (*)**

Use `lapply` function to load data from `urbanpop.xlsx` file, from *multiple sheets*, into a list called `pop_list`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-13}
**Data Manipulation: Merging**

Merge the 3 dataframes from `pop_list` using the `cbind()` function into a single dataframe `df`.

*Constraint:* Get rid of unnecessary columns "country" from the 2nd and 3rd datasets *inside* the `cbind` function when you're doing the merge.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-14}
**Data Cleaning: Regex & Formatting**

We've got data in a creepy format with "k" (thousands) and "M" (millions) instead of numbers.

1. Save the second column from `df` into `vec`.
2. Use `gsub()` to change "k" to "e3" and "M" to "e6" (scientific format).
3. Convert them to numeric and save into `vec_num`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-15}
**Data Cleaning: Automation**

Use the logic from the previous solution to apply this cleaning to the `df` dataframe and make the columns numeric.

```{r}
#| eval: false
# Your code here

```

:::