---
title: "05 Descriptive statistics | Part 1 (EDA)"
format: html
engine: knitr
---

::: {#exr-01}
**Data Import & Overview**

Load [`responses.csv`](class/responses.csv) to `df` and show the summary.
Then show all the column names and the first couple of rows, and discuss about the dataset.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-02}
**Categorical Data: Summary**

Give us a summary about Gender. How to make it useful?
What function should be used on `df$Gender`? 
Discuss why this variable may be useful in the research.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-03}
**Data Interpretation & Plotting**

Use `summary()` for spiders column.
Try to interpret each result. Then make plots with `plot()` for both.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-04}
**Subsetting by Groups**

Do the summary for spiders, but for different genders. Make plots as well.
For this you may'd like to create a subset. Discuss the result.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-06}
**Repeating Analysis**

Repeat the previous steps (summary, subsetting, and plotting) for Height and then for Weight.
Discuss results.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-07}
**Calculating Proportions**

Calculate the proportion of respondents who love music (values 4-5).

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-08}
**Feature Engineering: Faith**

Find how religious the youth is. Check variables God and Religion, create a new variable Faith which is equal to the sum of God and Religion divided by 2, rounded to the bottom with `floor()`. Check it with summary.
Check the percentage of unreligious (1-2) people who believe in God (4-5).

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-09}
**Advanced Summaries & sapply**

You'd like to show all mins, maxs, means and medians.
For this you have to work with numeric columns only to avoid errors.
Using `sapply` and function `is.numeric()` find all numeric columns. Then subset `df` with only those columns into `df_num`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-10}
**Applying Functions to Dataframes (\*)**

Continue with `df_num`. Use `sapply` with functions min/max/mean/median and `na.rm=T` and then explain the result.

```{r}
#| eval: false
# Your code here

```

:::