---
title: "06 Data Cleaning and Correlation Analysis"
format: html
engine: knitr
---

::: {#exr-01}
**Outlier Detection: IQR Rule**

Load the [`responses`](class/responses.csv) dataset into `df`. Use the **IQR Rule** to identify potential outliers for `Weight` and `Height`. 

::: {.callout-note}
Do not rewrite or modify the original data yet. You may save `df$Weight` into a variable `wght` and `df$Height` into `hght` for convenience. Discuss the results of the detection.
:::

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-02}
**Missing Values: Mean Imputation**

For the `wght` variable (Weight), fill all `NA` values with the **mean** value.

::: {.callout-tip}
You might want to use the `round()` function so the mean becomes an integer, matching the style of other observations. 
:::

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-03}
**Missing Values: Group-based Imputation**

1. Convert the `Gender` column into a **factor**.
2. Repeat the `NA` filling process for `hght` (Height), but this time do it **according to the gender** (use the mean height of males for missing male values, and vice versa).
3. Check if there is anything wrong or unusual in the process. 



```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-04}
**Handling Anomalies: Weight**

Find anomalies in `wght` using the **IQR Rule**. This time, rewrite (replace) the identified weight anomalies with the overall mean. 

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-05}
**Handling Anomalies: Height**

Find anomalies in `hght` using the **IQR Rule**. Rewrite these anomalies with the **mean of the corresponding gender**. 

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-06}
**Data Cleaning: Listwise Deletion**

Use the `na.omit()` function on the entire dataframe.

* How many observations are left?
* How should we interpret this result in the context of data loss? 



```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-07}
**Correlation: Physical Attributes**

Check the correlation between `df$Height` and `df$Weight`.

* Does the result look correct?
* If not, check the correlation between the cleaned variables `hght` and `wght` and discuss the difference. 



```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-08}
**Correlation: Beliefs**

Check the correlation between the variables `God` and `Religion`.

* To do this, you first need to handle `NA` values (either fill them or remove them).
* For this task, let's try removing them. 



```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-09}
**Correlation Matrix**

Create a correlation matrix for all variables in the dataset.

* *Step 1:* Filter the dataframe to include **only numeric** variables.
* *Step 2:* Use `na.omit()` to handle missing values.
* *Step 3:* Save the result into a matrix called `cormat`. 



```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-10}
**Visualization: Corrplot**

Install and load the `corrplot` library. Use the `corrplot()` function to visualize your `cormat`. Discuss the patterns you observe. 

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-11}
**Matrix Operations: Finding Extremes**

Find the **minimum value** in the `cormat` (the strongest negative correlation).

* Use the `which()` function with the parameter `arr.ind = TRUE` to find the exact row and column indices where the correlation equals the minimum value of the matrix.
* Discuss what these two variables represent. 



```{r}
#| eval: false
# Your code here

```

:::