---
title: "07 Data Manipulation with dplyr"
format: html
engine: knitr
---

::: {#exr-01}
**Data Import & Overview**

* Load necessary library `library(tidyverse)`.
* Load data from [`counties.csv`](class/counties.csv) using function `read_csv()` instead of the `read.csv()`.
* Show summary, first few rows and column names. 
* Discuss what is this dataset about.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-02}
**Pipe Notation**

Rewrite in a pipe notation following function:


`sort(unique(counties$state),decreasing=T)` 

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-03}
**Selecting Columns**

Select only the columns `state`, `county`, and `population` from counties using `select()` function.
Write into `counties_selected`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-04}
**Excluding Columns**

Select all columns except `population` and `region` from counties.
Use "-".

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-05}
**Arranging Data**

Arrange counties in descending order based on `private_work` percentage.
Using `arrange()` and `desc()` functions.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-06}
**Filtering Data**

Filter counties where `metro` is `Metro` and population is greater than 500,000 using `filter()`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-07}
**Creating Columns (Mutate)**

Create a new column `total_workforce` as the sum of `private_work` and `public_work` percentages.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-08}
**Multiple Filters**

Filter counties that belong to the `South` region and have a population under 50000.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-09}
**Calculating New Variables**

Create a new variable `self_employed_pop` as the number of self-employed individuals.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-10}
**Finding Top Observations**

Find the county with the highest percentage of public workers.
You may use `head()` to take only top observation.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-11}
**Filter and Sort**

Filter counties where more than 60% of the population is male and sort by population descending.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-12}
**Complex Pipeline**

For counties in a single pipe select variables `state`, `county`, `region`, `population`, `men`, `women`, `citizens`.
Then filter that there are more men than women and number of citizens is more than 10000 and the region is West or NorthEast.
Then arrange by citizens in descending order.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-13}
**Subsetting and Mutating Pipeline**

From counties select `state`, `county`, `population`, `men`, `women`, `self_employed`, add the proportion of men called `proportion_men`, take only counties with Proportion of men bigger than 60%, sort by `ProportionMen` in descending order and save into `counties_tmp`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-14}
**Modifying Existing Columns**

In `counties_tmp` remake `proportion_men` not as %, but as decimal with a function `mutate()`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-15}
**Debugging Code**

Why the following doesn't work? 

```{r}
#| eval: false
counties %>%
  select(state, county, population, men, women, self_employed)  %>%
  transmute(proportion_men=men/population*100) %>%
  filter(proportion_men>60) %>%
  arrange(desc(self_employed))
```

:::

---

::: {#exr-16}
**Using Transmute**

Use `transmute` over counties in a way to leave `state`, `county`, `population` and create `proportion_citizens`.
Save into `counties_tmp`.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-17}
**Proportions and Filtering**

Compute the proportion of self-employed individuals and filter only counties where it's above 10%, then sort by proportion descending.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-18}
**End-to-end Pipeline**

Create a pipeline that selects relevant columns, filters counties with at least 50000 population, creates a new variable for the male-to-female ratio, and sorts by this ratio.

```{r}
#| eval: false
# Your code here

```

:::

---

::: {#exr-19}
**Advanced Pipeline**

Select `state`, `county`, `population`, `private_work`, `public_work`, `self_employed`, compute into total_employment the total percentage of employment by summing up private, public, and self-employment percentages, then filter only counties where it's below 98% from Texas and arrange by descending order.

```{r}
#| eval: false
# Your code here

```

:::