---
title: "08 Data Aggregation and Grouping with dplyr"
format: html
engine: knitr
---

::: {.callout-important}
* Load necessary library `library(tidyverse)`.
* Load data from [`counties.csv`](class/counties.csv).
:::

::: {#exr-01}
**Counting Categories**

Count the number of counties in each metro/non-metro category.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-02}
**Conditional Counting**

Count the number of counties in each state where the average household income is above 60,000.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-03}
**Weighted Counting**

Count the number of counties in each state, weighted by the number of employed people, and sort in descending order.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-04}
**Multi-group Counting**

Count the number of metro and non-metro counties per state.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-05}
**Average Percentage (Two Ways)**

Find the average percentage of white people throughout all counties.
Propose 2 approaches, one with `mean()` and another with `summarize()`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-06}
**Conditional Average**

Find the average percentage of white people throughout all counties, which have income more than 70,000.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-07}
**Finding the Maximum**

Find the state with the maximum average household income.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-08}
**Finding the Minimum Amount**

Find the county with the lowest amount of people working in the public sector.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-09}
**Summarizing Totals**

Compute the total number of unemployed people in each state and sort in descending order.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-10}
**Mean and Median**

Find the average and median number of people per county working in private companies.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-11}
**Calculating Percentages**

Calculate the percentage of the population that works in construction per state.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-12}
**Grouped Maximum**

Find the state with the highest average unemployment rate among metro counties.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-13}
**Multiple Metrics**

Compute the total number of counties per state and the average number of people walking to work per state.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-14}
**Highest Proportion**

Find the state with the highest proportion of people working in self-employment.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-15}
**Proportion of Citizens**

Compute the proportion of people in each state who are U.S. citizens.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-16}
**Complex Pipeline**

For the `counties` dataset do the following:

1. Select variables `state`, `population`, `white`, `black`, `hispanic`, `native`, `asian`.
2. Filter for counties where there are more than 20% of black persons, more than 10% of hispanic persons, and the population is bigger than 150,000.
3. Show the total amount (sum) of black, hispanic, native, and asian people by state.

```{r}
#| eval: false
# Your code here
```
:::