---
title: "09 Tidying Data and Joins (tidyr & dplyr)"
format: html
engine: knitr
---

::: {#exr-01}
**Pivoting Wider: Basic**

Convert the counties dataset into a wide format where the columns are `county`, `state`, `metro` or `non-metro` (after pivoting) and values are `population`. 
Do it with `pivot_wider()`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-02}
**Pivoting Wider: Aggregation**

Convert the counties dataset into a wide format where for each state the population is calculated by metro and non-metro. 
So we'll have columns `state`, `metro`, `nonmetro` and 50 rows for each state.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-03}
**Pivoting Wider: States as Columns**

Convert the counties dataset into a wide format where each state becomes a column and the values are the total population in that state. 
Save into `counties_pop_wide`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-04}
**Pivoting Longer**

Convert the wide-format data back into a long format where the state names become a column again from `counties_pop_wide`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-05}
**Summary and Pivoting**

Create a summary table showing the average income in each metro and non-metro area per region, then pivot it wider by metro and save into `counties_reg_metro`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-06}
**Joining Data: left_join**

Create another table by region for average income calculated correctly, save it into `counties_reg_tot` and connect it to `counties_reg_metro` with the `left_join()` and resave the result as the `counties_reg_tot`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-07}
**Complex Pipeline: Population & Income (Part 1)**

Create a total population by region and metro, make it wider to have metro and nonmetro columns, save into `counties_reg_metro_pop`. 
Then do the same but just by region and save into `counties_metro_pop`. 

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-08}
**Complex Pipeline: Population & Income (Part 2)**

To `counties_reg_metro_pop` join the `counties_metro_pop` and calculate the share of metro population in total population, same for nonmetro. 
Then join the `counties_reg_tot` by region and give suffixes `_perc` and `_inc` for percentages and income.
At the last stage, compare `mean_inc` with `mean_inc_new`, which is calculated as mean income for metro and non-metro multiplied by the percentage of population in metro and non-metro area. Are they the same? Why?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-09}
**Joining External Data**

Perform a left join between the `counties` dataset and a dataset containing state abbreviations (e.g., `state.abb` and `state.name` from base R), ensuring that no counties are lost in the process.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-10}
**Cross Tabulation & Pivoting**

Create a table with percentage of white, black, and other people (columns) and metro / non-metro county (rows).

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-11}
**Calculations and Joins**

Find the population density for each state and the average income for each state, then join them.
Check the correlation between two.

```{r}
#| eval: false
# Your code here
```
:::