---
title: "10 Data Manipulation with IMDb Dataset"
format: html
engine: knitr
---

::: {#exr-01}
**Data Loading and Joining**

Load [`ImdbTitleBasics`](https://drive.google.com/file/d/1FzYor9a8oY-6GKQAkbUGF4xO87nAlVmT/view?usp=drive_link) and [`ImdbTitleRatings`](https://drive.google.com/file/d/1r4mHgnORXQVPs6ugUDMZoaY9vT4rc2fG/view?usp=drive_link) into `imdb_1` and `imdb_2`. Check column names and first few rows. 
Join `imdb_1` and `imdb_2` into `imdb`, and remove `imdb_1` and `imdb_2`.

```{r}
#| eval: false
library(tidyverse)
library(googledrive)

# 1. Specify the full URLs
link_1 <- "https://drive.google.com/file/d/1FzYor9a8oY-6GKQAkbUGF4xO87nAlVmT/view?usp=drive_link"
link_2 <- "https://drive.google.com/file/d/1r4mHgnORXQVPs6ugUDMZoaY9vT4rc2fG/view?usp=drive_link"

# (Optional) If the files are publicly accessible via a link, this command 
# will allow you to download them WITHOUT logging in (R will not open a browser window asking you to log in)
drive_deauth()

# 2. Upload the files. as_id() will automatically retrieve the required ID from the link
drive_download(as_id(link_1), path = "ImdbTitleBasics.csv", overwrite = TRUE)
drive_download(as_id(link_2), path = "ImdbTitleRatings.csv", overwrite = TRUE)

# 3. Reading local files
imdb_1 <- read_csv("ImdbTitleBasics.csv")
imdb_2 <- read_csv("ImdbTitleRatings.csv")
```
:::

::: {.callout-tip collapse="true"}
## Or alternativly
```{r}
#| eval: false
library(tidyverse)
library(googledrive)

drive_deauth()

# Creating vectors with links and filenames
links <- c(
  "https://drive.google.com/file/d/1FzYor9a8oY-6GKQAkbUGF4xO87nAlVmT/view?usp=drive_link",
  "https://drive.google.com/file/d/1r4mHgnORXQVPs6ugUDMZoaY9vT4rc2fG/view?usp=drive_link"
)

filenames <- c("ImdbTitleBasics.csv", "ImdbTitleRatings.csv")

# We automatically iterate through both vectors and download each file
walk2(links, filenames, ~drive_download(as_id(.x), path = .y, overwrite = TRUE))

# Reading Files
imdb_1 <- read_csv(filenames[1])
imdb_2 <- read_csv(filenames[2])
```
:::



---

::: {#exr-02}
**Data Cleaning**

Clean the `imdb` a little bit. First check the unique genres in the dataset, discuss. Then remove all the movies where the "Adult" in genres (use `str_detect()` for it) and also remake `startYear`, `endYear` and `runtimeMinutes` as numeric.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-03}
**Top Movies by Runtime**

Show top-2 movies by the runtime in each genre, checking only titles with 1 genre. 
*Tip:* If there are 2+ genres, what symbol will definitely be in genres?
*Tip 2:* You may need function `slice_head()` to take 2 titles from each group.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-04}
**Sorting by Decades**

Sort in descending order title average scores for each decade, taking only titles with 1k+ votes.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-05}
**Counting by Format**

For movies and tvseries make a table 1x2 with amount of movies and TV series.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-06}
**Most Voted per Year**

Find the most voted title for each year.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-07}
**Rating Distributions**

Compare the rating distribution of movies and TV series. Calculate Mean, median and standard deviation.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-09}
**Separating Rows by Genres**

Use `separate_rows()` on genres with `sep=","` to divide titles with multiple genres into multiple rows where each observation may appear multiple times, but with different genres. 
Check the amount of observations in this case. Compare with the amount of observations of just `imdb` and the case where you leave only titles with a single genre. 
Now check unique genres for `imdb` with separated rows and compare.

```{r}
#| eval: false
# Your code here
```
:::