---
title: "11 Data Visualization: Scatter Plots, Histograms, and Bar Charts"
format: html
engine: knitr
---

::: {#exr-01}
**Scatter plot with log GDP and continent colors**

Create a scatter plot of logarithmized GDP per capita vs Life Expectancy, colored by continent. Use point size 2.

```{r}
#| eval: false
library(tidyverse)
library(gapminder)

gpmind <- gapminder

# Your code here
```
:::

---

::: {#exr-02}
**Scatter plot with population-based size**

Create a scatter plot of log GDP per capita vs Life Expectancy with color by continent, size based on population, and transparency 0.6.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-03}
**Scatter plot for 2007 with minimal theme**

Create a scatter plot with log GDP per capita and Life Expectancy colored by continent for year 2007 only. Use point size 5 and add `theme_minimal()`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-04}
**Scatter plot with smooth trend lines**

Create a scatter plot with smooth trend lines per continent. Use both `geom_point()` (alpha=0.6, size=3) and `geom_smooth(se=F)` with line width 3.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-05}
**Scatter plot with country labels**

Create a scatter plot with country labels instead of dots. Use `geom_text()` with log GDP per capita on x-axis and Life Expectancy on y-axis, colored by continent.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-06}
**Annotated scatter plot with year labels**

Calculate average life expectancy and average GDP per capita by year and continent. Create a scatter plot with year labels using a combination of `geom_point()` and `geom_text(vjust=-0.5)`. Add `theme_minimal()`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-07}
**Histogram of Life Expectancy in 2007**

Create a histogram of Life Expectancy in 2007 with `binwidth=3`, filled with `"lightblue3"` and black border.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-08}
**Histogram filled by continent**

Create a histogram of Life Expectancy filled by continent with `binwidth=1`, black border, and transparency 0.7.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-09}
**Density plot by continent**

Create a density plot of Life Expectancy in 2007 by continent with transparency 0.3.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-10}
**Bar chart of countries per continent**

Create a bar chart of the number of countries per continent in 2007. Try to do it using both `geom_col()` and `geom_bar()`. In which case did you succeed?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-11}
**Bar chart: weighted vs simple mean**

Create a bar chart of average life expectancy by continent in 2007, comparing weighted average (`sum(lifeExp*pop)/sum(pop)`) and simple mean. Use `pivot_longer()` and `position = "dodge"`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-12}
**Median GDP bar chart**

Create a bar chart of median GDP per capita by continent in 2007, filled with `"lightblue3"`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-13}
**Grouped bar chart by year**

Create a bar chart of maximum life expectancy by continent for years 1997, 2002, and 2007. Use `position="dodge"` and color by year (as factor).

```{r}
#| eval: false
# Your code here
```
:::

