---
title: "13 Data Visualization: Themes, Fonts, and Custom Styling"
format: html
engine: knitr
---

::: {.callout-tip}
## Links

- [MetBrewer 🎨](https://github.com/BlakeRMills/MetBrewer)
- [MoMAColors 🎨](https://github.com/BlakeRMills/MoMAColors)
- [ColorBrewer 2.0](https://colorbrewer2.org/)
- [CARTO Colors](https://carto.com/carto-colors/)
- [rcartocolor](https://jakubnowosad.com/rcartocolor/)

:::

::: {#exr-01}
**Base line chart**

Build a line chart of weighted average GDP per capita over time by continent. Save the result into an object `ggp`.

```{r}
#| eval: false
library(tidyverse)
library(gapminder)
library(showtext)

font_add_google("Pacifico", "cute_font")
showtext_auto()

gpmind <- gapminder

# Your code here
```
:::

---

::: {#exr-02}
**Add labels**

Add a meaningful title, x-axis label, y-axis label, and caption to `ggp`. Resave into `ggp`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-03}
**Apply a theme with a custom font**

Add `theme_minimal()` to `ggp`. Then try passing `base_family = "cute_font"` as a parameter to `theme_minimal()`. Resave `ggp` with the font-based theme.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-04}
**Custom color palette**

Create a vector `palette_cute` with the colors `"#FF69B4"`, `"#C71585"`, `"#8B008B"`, `"#FFA07A"`, `"#DC143C"`. Apply it to `ggp` using `scale_color_manual()`, assigning each color to its continent by name. Resave `ggp`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-05}
**Plot background color**

Change the plot background color of `ggp` to light pink `"#FFF0F5"` using `theme(plot.background = ...)`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-06}
**Panel background color**

Change the panel (chart area) background color to soft peach `"#FFE4E1"`. Keep the plot background from the previous exercise.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-07}
**Legend background**

Modify the legend background and legend key colors to baby pink `"#FFD1DC"`. Keep all previous styling.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-08}
**Legend text styling**

Customize the legend text: size 14, color `"#8B008B"`. Keep all previous styling.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-09}
**Title styling**

Style the plot title: bold, size 20, color `"#DC143C"`. Keep all previous styling.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-10}
**Axis title styling**

Customize axis titles: size 16, color `"#C71585"`. Keep all previous styling.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-11}
**Axis text styling**

Change the axis text (tick labels) color to deep pink `"#FF1493"` and size to 14. Keep all previous styling.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-12}
**Create a reusable custom theme**

Combine all the styling from previous exercises into a single `theme_cute` object. Apply it to `ggp` and adjust the title to span two lines with `\n`. Use larger font sizes (legend text 20, title 28, axis titles 24, axis text 20) and add caption and legend title styling.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-13}
**Build the KSE palette**

Create a vector `palette_kse` with 4 brand colors of KSE: `"#003964"`, `"#00bbce"`, `"#A7C539"`, `"#D33E2C"`. Apply it to the weighted average GDP line chart using `scale_color_manual()`.

```{r}
#| eval: false
# Your code here
```
:::
