---
title: "14 Introduction to Linear Regression"
format: html
engine: knitr
---

::: {.panel-tabset}

### Regression Analysis, Part 1

[Link, Part 1](https://aranaur.quarto.pub/ds24lec09)

```{=html}
<iframe class="slide-deck" src="https://aranaur.quarto.pub/ds24lec09/#/title-slide" height="420" width="747" style="border: 1px solid #581B39;"></iframe>
```

### Regression Analysis, Part 2

[Link, Part 2](https://aranaur.quarto.pub/ds24lec10)

```{=html}
<iframe class="slide-deck" src="https://aranaur.quarto.pub/ds24lec10/#/title-slide" height="420" width="747" style="border: 1px solid #581B39;"></iframe>
```

:::


```{r}
#| eval: false
#| label: setup
library(tidyverse)
library(gapminder)
library(Metrics)

gpmind <- gapminder %>%
  select(country, continent, year, lifeExp, pop, gdpPercap)
```

---

## Part 1: What is a linear regression?

A linear regression is a **rule** that predicts one variable from another.
The rule looks like: `Y = a + b * X`, where `a` is the intercept and `b` is the slope.
We choose the rule that minimizes the total squared prediction error.

::: {#exr-01}
**Explore the data**

Print a summary of `gpmind`. What is a single observation here? What are `lifeExp` and `gdpPercap`? Which variable would you predict and why?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-02}
**Scatter plot: raw GDP vs Life Expectancy**

Build a scatter plot of `gdpPercap` on X and `lifeExp` on Y. Color points by continent and set `alpha = 0.4`. Does the relationship look linear?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-03}
**Scatter plot: log GDP vs Life Expectancy**

Build the same scatter plot but with `log(gdpPercap)` on X. Does it look more linear now? What kind of formula could describe this line?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-04}
**Fit a simple linear model**

Fit a linear model of `lifeExp` on `log(gdpPercap)`. Save it as `lm1` and print it to see the coefficients. What do the two numbers (intercept and coefficient) mean?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-05}
**Interpret the summary**

Print `summary(lm1)`. Focus only on the Coefficients table. If `log(gdpPercap)` increases by 1, what happens to `lifeExp`? Is this a causal statement?

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 2: Fitted values, residuals, and what the model actually does

The model produces for each observation:

- a **fitted value** — what it predicts
- a **residual** — how wrong it was (`actual − fitted`)

::: {#exr-06}
**Extract fitted values and residuals**

Extract the fitted values and residuals from `lm1` (they are elements inside the `lm1` object). Print the first 6 of each using `head()`. Verify by hand that `residual = lifeExp − fitted value` for the first observation.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-07}
**Verify residuals formally**

Verify for **all** observations that `residual == lifeExp − fitted.values` (round to 3 decimals). Use `sum()` and `==` to count how many match.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-08}
**Fitted vs actual plot**

Plot fitted values on X and actual `lifeExp` on Y. Add `geom_abline()` in blue with `linewidth = 2`. What does a point far from the diagonal line mean?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-09}
**Distribution of residuals**

Plot the distribution of residuals using `geom_density()`. First convert residuals from `lm1` to a data frame. What shape would we ideally want to see and why?

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 3: How good is the model? R-squared

**R-squared** answers: how much of the variation in Y does our model explain?
It ranges from 0 (explains nothing) to 1 (explains everything).

::: {#exr-10}
**Build R-squared by hand**

Save `summary(lm1)` into `smr1`. Then:

1. Compute `TSS` = sum of squared deviations of `lifeExp` from its mean.
2. Compute `RSS` = sum of squared residuals from `lm1`.
3. Compute `R-squared = 1 − RSS/TSS`.

Verify it matches `smr1$r.squared`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-11}
**R-squared as squared correlation**

Verify that R-squared equals the squared Pearson correlation between fitted values and actual `lifeExp`. Compare to `smr1$r.squared`.

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 4: Are the coefficients significant?

A p-value tells us: if the true coefficient were zero, how likely is it to observe a coefficient this large just by random sampling? Convention: p-value < 0.05 is statistically significant.

::: {#exr-12}
**Extract p-values**

Extract the full coefficients table from `smr1`. Then extract only the p-values (column 4) and check which are below 0.05.

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 5: Multiple regression — adding more predictors

We can extend the rule: `Y = a + b1*X1 + b2*X2 + ...`
Each coefficient means: the effect of X1 **holding all other X constant**.

::: {#exr-13}
**Add year as predictor**

Fit `lm2 <- lm(lifeExp ~ log(gdpPercap) + year, data = gpmind)`. Print `summary(lm2)`. How did R-squared change? Interpret the coefficient on `year`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-14}
**Add continent as predictor**

Fit `lm3 <- lm(lifeExp ~ log(gdpPercap) + year + continent, data = gpmind)`. Print `summary(lm3)`. What happened to R-squared? Why is one continent missing from the table?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-15}
**Compare three models**

Print R-squared and adjusted R-squared for `lm1`, `lm2`, and `lm3`. Why is adjusted R-squared useful when adding predictors?

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 6: RMSE — measuring error in interpretable units

RMSE (Root Mean Squared Error) measures error in the same units as Y.
Roughly: on average, predictions are off by RMSE units.

::: {#exr-16}
**Compute RMSE manually and with Metrics**

Compute RMSE for `lm1` manually as `sqrt(mean(residuals^2))`. Then compute it using `Metrics::rmse()`. Do they match?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-17}
**Compare RMSE across models**

Compute RMSE for all three models using `Metrics::rmse()`. Which model has the lowest error?

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 7: Making predictions

::: {#exr-18}
**Predict life expectancy for Ukraine (lm1)**

Use `lm1` to predict life expectancy for Ukraine assuming average annual income = `12 * 600` USD. Create a one-row data frame with that `gdpPercap`, then use `predict()`.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-19}
**Predict with multiple predictors (lm3)**

Use `lm3` to predict life expectancy for Ukraine in 2007. Create a one-row data frame with `gdpPercap = 7200`, `year = 2007`, `continent = "Europe"` (as a factor with levels from `gpmind`). Is the `lm3` prediction better or worse than `lm1`?

```{r}
#| eval: false
# Your code here
```
:::

---

## Part 8: Train / test split — does the model generalize?

A train/test split lets us check if the model generalizes to **new data**.

::: {#exr-20}
**Split into train and test**

Split `gpmind` into 75% training and 25% test sets using `runif()`. Use `set.seed(42)`. Call them `gp_train` and `gp_test`. Print the number of rows in each.

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-21}
**Train RMSE vs test RMSE**

Train `lm1` again but only on `gp_train`. Call it `lm_tr`. Compute RMSE on the training set and RMSE on the test set. Are they close? What would a big gap mean?

```{r}
#| eval: false
# Your code here
```
:::

---

::: {#exr-22}
**Plot predictions on the test set**

Plot predicted vs actual `lifeExp` for the **test set only**. Add `geom_abline()` in blue. Write a comment about what you observe.

```{r}
#| eval: false
# Your code here
```
:::
